---
title: "Lab 6 Solutions"
author: ""
date: ""
output: 
  html_document:
    css: lab6.css
    toc: true
    number_sections: false
    toc_float: true
    df_print: paged
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE,
                      message = FALSE, warning = FALSE, comment = NA,
                      fig.width=8, fig.height=6)
```

## Introduction

On May 20, 2013 an extremely powerful and devastating tornado struck
Moore, Oklahoma, a city just south of Oklahoma City. The tornado was over one
mile wide and had a maximum wind speed of 210 mph. This area is a known hotbed 
for tornadic activity. In fact, the 2013 Moore tornado followed a very similar 
path as two other large tornadoes that hit the area in 1999 and 2003, 
respectively.

In this lab we'll scrape and wrangle tornado data provided by the Tornado
History Project.

## Packages

```{r}
library(tidyverse)
library(rvest)
library(janitor)
library(sf)
library(mapview)
```

## Scrape data

#### Step 1

<div class="bkgemph">
Scrape data from http://www.tornadohistoryproject.com to get all
tornadoes recorded in Oklahoma in 1998. Your result should be a data frame 
object named `tor_1998`. Hint: make use of selector gadget and
`html_table()`. In step 2 you can focus on cleaning up the data frame.
</div>

```{r tornadoes_1998}
url <- "http://www.tornadohistoryproject.com/tornado/Oklahoma/1998/table"
tor_1998_html <- read_html(url)

tor_1998 <- tor_1998_html %>% 
  html_nodes("#results") %>% 
  html_table() %>% 
  .[[1]]
```

#### Step 2

<div class="bkgemph">
Clean up `tor_1998` by removing the rows with header information, fix the
variable names with `janitor::clean_names()`, and only keep columns
`spc_number:lift_lon`.
</div>

```{r clean_tornadoes}
names(tor_1998) <- tor_1998[1, ]

tor_1988_clean <- tor_1998 %>% 
  filter(Date != "Date") %>% 
  clean_names() %>% 
  select(spc_number:lift_lon) %>% 
  as_tibble()
```

#### Step 3

<div class="bkgemph">
Use your code from steps 1 and 2 to now create a function named 
`get_tornadoes()`. The function should have a single argument, `year`. Your
function should be able to scrape the tornado data for any year from 1998 to 
2017 and output the cleaned data frame.
</div>

```{r get_tornadoes}
get_tornadoes <- function(year) {
  base_url <- "http://www.tornadohistoryproject.com/tornado/Oklahoma/"
  url <- str_c(base_url, year, "/table")
  
  tor_html <- read_html(url)
  
  tor <- tor_html %>% 
    html_nodes("#results") %>% 
    html_table() %>% 
    .[[1]]  
  
  names(tor) <- tor[1, ]

  tor %>% 
    filter(Date != "Date") %>% 
    janitor::clean_names() %>% 
    select(date:lift_lon) %>% 
    as_tibble()
}
```

#### Step 4

<div class="bkgemph">
Use your function `get_tornadoes()` to scrape all the tornado data from 1998 to
2017. Choose a `map` function that will do this with one line of code and 
result in a single data frame. Name the resulting data frame `ok_tornadoes`.
</div>

```{r get_all_tornadoes}
ok_tornadoes <- map_df(1998:2017, get_tornadoes)
```

## Visualize data

<div class="bkgemph">
Use `ok_tornadoes` to create a visualization. You may need to further
clean up the variables as there are special characters in the data. For example,
`fujita` has some `?` present.
</div>

**Answers will vary**

```{r create_sf_object}
# create a simple feature object, use touch data, EPSG = 4267
ok_sf <- ok_tornadoes %>% 
  select(-contains("lift")) %>% 
  mutate(fujita = str_replace_all(fujita, pattern = "\\?", replacement = "0"),
         fujita = factor(fujita)) %>% 
  mutate_at(.vars = vars(contains("touch")), as.numeric) %>% 
  st_as_sf(coords = c("touch_lon", "touch_lat")) %>% 
  st_set_crs(value = 4267)
```

```{r visualize_tornadoes}
# use mapview to create visualization
mapviewOptions(legend.pos = "bottomleft")
mapview(ok_sf, zcol = "fujita", layer.name = "Tornado Fujita Scale",
        col.regions = c("#feedde", "#fdd0a2", "#fdae6b", "#fd8d3c", "#e6550d",
                        "#a63603"),
        cex = 4)
```

## References

1. Tornado History Project: Maps and Statistics . (2020). 
   Tornadohistoryproject.com. Retrieved 1 June 2020, from
   http://www.tornadohistoryproject.com/